---
title: "HW8"
author: "phil newsome"
echo: false
format: pdf
---

```{r}
library(here)
library(cmdstanr)
library(dplyr)
library(brms)
library(modelsummary)
library(posterior)
library(bayesplot)
```

# Research Question

Does fear recognition accuracy predict prenatal-postpartum change in amygdala responses to fear faces in first-time fathers?

# Variables

-   'hit_rate' : unbiased hit-rate (raw accuracy and differential accuracy combined)

-   'lamyg_mean_change' : change in left amygdala activation to fear\>rest contrast

-   'ramyg_mean_change' : change in right amygdala activation to fear\>rest contrast

```{r}
# data import
faces <- read.csv("faces_bayesian.csv")
faces <- na.omit(faces)

```

Table \@tbl-summ-var shows summary statistics

```{r}
# variable summary 
#| label: tbl-summ-var
#| tbl-cap: Summary Statistics 
datasummary(
  hit_rate + ramyg_mean_change + lamyg_mean_change ~ 
  N + Mean + SD + Min + Max + Histogram,
  data = faces
)

```

# Model (w. left amygdala)

Let $Y$ = hit_rate, $G$ = lamyg_mean_change

model (left) $$
\begin{aligned}
    Y_i & \sim N(\mu_i, \sigma) \\
    \mu_i & = \beta_0 + \beta_1 G_i
\end{aligned}
$$

priors: $$
\begin{aligned}
    \beta_0 & \sim N(0.6, .3) \\
    \beta_1 & \sim N(0, 0.2) \\
    \sigma & \sim N^+(0, 1)
\end{aligned}
$$

# Analysis

```{r}
#| include: false
mod1 <- cmdstan_model(here("faces_bayesian_model.stan"))

# 1. Form the data list for Stan
stan_dat <- list(
  N = nrow(faces),  # Total number of observations
  hit_rate = faces$hit_rate,  # The dependent variable (hit_rate)
  lamyg_mean_change = faces$lamyg_mean_change  # The predictor variable (lamyg_mean_change)
)

# 3. Run Stan (fit the model)
fit <- mod1$sample(
  data = stan_dat,
  seed = 1234,  # for reproducibility
  iter_warmup = 1000,  # Number of warmup iterations
  iter_sampling = 2000,  # Number of sampling iterations
  chains = 4  # Number of chains
)

```

# Results

As shown in the rank histogram in \@fig-rank-hist-fit below, the chains mixed well.

```{r}
#| label: fig-rank-hist-fit
#| fig-cap: Rank histogram of the posterior distributions of model parameters.
as_draws(fit) |>
    mcmc_rank_hist(pars = c("beta0", "beta1", "sigma"))
```

```{r}
as_draws(fit) |>
  mcmc_dens(pars = c("beta0", "beta1", "sigma"))

```

```{r}
as_draws(fit) |>
  mcmc_trace(pars = c("beta0", "beta1", "sigma"))

```

\@tbl-summ-fit shows the posterior distributions

```{r}
#| label: tbl-summ-fit
#| tbl-cap: Posterior summary of the model parameters.

draws <- as_draws(fit)

summ_fit <- fit$draws() |>
    subset_draws(variable = c("beta0", "beta1", "sigma")) |>
    summarise_draws()
knitr::kable(summ_fit, digits = 2)
```

Evidence for a positive relationships between fear recognition accuracy and left amygdala response change is inconclusive, with a posterior mean of 0.03 and a 90% CI of \[-0.23, 0.29\].
